Pretty darn good control: when are approximate solutions better than approximate models

### 2023-01-23 manuscript/manuscript.Rmd
mini
state space figure

# rename some columns for figure purposes:
colnames(msy) <- c('t', 'rep', 'action', 'reward', 'X', 'Y', 'Z')
colnames(ppo) <- c('t', 'rep', 'action', 'reward', 'X', 'Y', 'Z') 
colnames(escapement) <- c('t', 'rep', 'action', 'reward', 'X', 'Y', 'Z')

@@

best_e <- 0.62 # optimal constant escapement, computed from ../data/escapement.csv.gz
esc_sim <- escapement |> 
    rename(escapement = action) |> # data had a naming problem
    filter(escapement == best_e) |> 
    pivot_longer(c("X", "Y", "Z"),
        names_to = "species", values_to = "abundance") |>
    mutate(abundance = abundance + 1)

@@

Our results suggest that DRL can find solutions that are more sustainable *and* more profitable than any constant mortality strategy.
Indeed, DRL finds solutions in which, for certain periods of time, no harvest is performed on the system.
These periods effectively allow the system to, in a sense, "reset" itself.
They lead to a distinctive oscillating behavior of the system state over time, with periods of relative depletion followed by periods of growth. 
Figure \ref{fig:tbd} exemplifies this point by comparing one times series of the system state obtained using each of these policies.

```{r timeseries_fig, fig.cap="TBD.", out.width = "6.5in"}
policy_labels <- list("msy" = "Opt. Const. Mortality", "tac"= "80% Opt. Const. Mortality", "ppo"= "DRL")
policy_labeller <- function(variable,value){
  return(policy_labels[value])
}
msy_tac_ppo |> 
    group_by(rep, policy) |>
    filter(rep == 0) |>
    ggplot(aes(t, abundance, col = species)) + 
        geom_line()  + 
        scale_color_manual(values=c('palegreen3', 'coral1', 'deepskyblue4')) +
        facet_wrap(~factor(policy, levels = c("msy", "tac", "ppo")), labeller = policy_labeller ) +
        labs(y = "abundance", x = "time") +
        theme(aspect.ratio = 2/3)
```

@@

(p1 / p2)
```

It is well known that following MSY policies can lead to over-fishing.
Our results reproduce this behavior and, furthermore, warn that using a TAC policy instead might not be enough to gain sustainability.
Figure TBD-A summarizes episode length distribution for MSY, TAC and DRL.
While the DRL-derived policy does not lead to extinctions, both MSY and TAC do so, and thus their episode mean length is lower than the maximal episode length (in our case set at 200 years).
This difference is reflected on the profits earned by the agent, as shown in Fig. TBD-B.

>>>>

# rename some columns for figure purposes:
colnames(msy) <- c('t', 'rep', 'action', 'reward', 'X', 'Y', 'Z')
colnames(ppo) <- c('t', 'rep', 'action', 'reward', 'X', 'Y', 'Z') 
colnames(escapement) <- c('t', 'rep', 'esc_level', 'reward', 'X', 'Y', 'Z') 
opt_escapement <- escapement |> filter(esc_level == 0.62)
head(opt_escapement)

@@

best_e <- 0.62 # optimal constant escapement, computed from ../data/escapement.csv.gz
esc_sim <- escapement |> 
    pivot_longer(c("X", "Y", "Z"),
        names_to = "species", values_to = "abundance") |>
    mutate(abundance = abundance + 1)

@@

Our results suggest that DRL can find solutions that are more sustainable *and* more profitable than any constant mortality strategy.
Indeed, DRL finds solutions in which, for certain periods of time, no harvest is performed on the system.
These periods effectively allow the system to, in a sense, "reset" itself.
They lead to a distinctive oscillating behavior of the system state over time, with periods of relative depletion followed by periods of growth. 
Figure TBD exemplifies this point by comparing one times series of the system state obtained using each of these policies.

```{r timeseries, fig.cap="TBD.", out.width = "6.5in"}
policy_labels <- list("msy" = "Opt. Const. Mortality", "tac"= "80% Opt. Const. Mortality", "ppo"= "DRL")
policy_labeller <- function(variable,value){
  return(policy_labels[value])
}
msy_tac_ppo |> 
    group_by(rep, policy) |>
    filter(rep == 0) |>
    ggplot(aes(t, abundance, col = species)) + 
        geom_line()  + 
        scale_color_manual(values=c('palegreen3', 'coral1', 'deepskyblue4')) +
        facet_wrap(~factor(policy, levels = c("msy", "tac", "ppo")), labeller = policy_labeller ) +
        labs(y = "abundance", x = "time") +
        theme(aspect.ratio = 2/3)
```

@@

(p1 / p2) + plot_annotation(tag_levels = 'A')
```

```{r state-space}
i_tac <- which.min(abs(actions - best_action * 0.8))

# color pallette limits
library(RColorBrewer)
myPalette <- colorRampPalette(rev(brewer.pal(11, "Spectral")))
sc <- scale_colour_gradientn(colours = myPalette(100), limits=c(0,1))

pmsy <- msy |>
  mutate(X = X+1, Y = Y+1, Z = Z+1, harvest = action * X) |>
  filter(action == best_action) |>
  ggplot(aes(X, Z, color = harvest)) +
  geom_point() + sc +
  theme(aspect.ratio = 1/1)

ptac <- msy |>
  mutate(X = X+1, Y = Y+1, Z = Z+1, harvest = action * X) |>
  filter(action == actions[[i_tac]]) |>
  ggplot(aes(X, Z, color = harvest)) +
  geom_point() + sc +
  theme(aspect.ratio = 1/1)

pppo <- ppo |>
  mutate(X = X+1, Y = Y+1, Z = Z+1, harvest = action * X) |>
  ggplot(aes(X, Z, color = harvest)) +
  geom_point() + sc +
  theme(aspect.ratio = 1/1)

pesc <- opt_escapement |>
  mutate(X = X+1, Y = Y+1, Z = Z+1, 
        harvest = case_when(
          X - esc_level > 0 ~ X - esc_level,
          X - esc_level <= 0 ~  0)
         ) |> 
  ggplot(aes(X, Z, color = harvest)) +
  geom_point() + sc +
  theme(aspect.ratio = 1/1)

(
  (pmsy | ptac) / (pppo | pesc) +
  plot_layout(guides = "collect") & theme(legend.position = 'right')
) + 
  plot_annotation(tag_levels = c('A'))
```

**Probably more for the discussion**
It is well known that following MSY policies can lead to over-fishing.
Our results reproduce this behavior and, furthermore, warn that using a TAC policy instead might not be enough to gain sustainability.
Figure \ref{rew_len_fig}-A summarizes episode length distribution for MSY, TAC and DRL.
While the DRL-derived policy does not lead to extinctions, both MSY and TAC do so, and thus their episode mean length is lower than the maximal episode length (in our case set at 200 years).
This difference is reflected on the profits earned by the agent, as shown in Fig. TBD-B.
